---
title: "Replication of Problems 1 and 2 of Tversky & Kahneman (1981, *Science*)"
author: "Your Name (your.email@stanford.edu)"
date: today
date-format: long
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: false
    embed-resources: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
execute:
  warning: false
  message: true
---

<!--
PSYCH 251 replication report template.

Every project uses this structure so that reports are comparable across the class. They
become public supplementary material, so write for a reader who has the original paper but
not your repository.

The Introduction and Methods are written, critiqued, and preregistered BEFORE data collection.
Render this file at every stage: before you have data the Results section says so and skips
itself; during Pilot A it runs your analysis on pilot data, which is the point of Pilot A.

Grey "Guidance" boxes are instructions. Delete each one once you have written that section.
The code is wired to the template's demo experiment (the framing problem). Replace the demo
specifics with your study's as you go: the experiment id, the original study's numbers, the
exclusion rule, and the confirmatory analysis.
-->

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(jsonlite)

# Must match EXPERIMENT.id in experiment.js, and the --experiment you passed to the export.
experiment_id <- "framing-demo"

raw_dir <- file.path("..", "data", "raw", experiment_id)
have_data <- all(file.exists(file.path(raw_dir, c("participants.csv", "trials.csv"))))
```

```{r}
#| label: original-study
#| include: false
# The original study's numbers, entered once and used by the power analysis and by the
# replication assessment. Replace these with your original study's values, and say in a
# comment exactly where in the paper each one comes from.
#
# Tversky & Kahneman (1981), Problem 1 (gain frame, N = 152): 72% chose the certain program,
# 28% the risky one. Problem 2 (loss frame, N = 155): 22% certain, 78% risky. The paper
# reports percentages, so the counts below are reconstructed from them.
original <- tibble(
  condition = c("gain", "loss"),
  n         = c(152, 155),
  risky     = round(c(0.28 * 152, 0.78 * 155))
)
original_p_gain <- original$risky[original$condition == "gain"] / original$n[original$condition == "gain"]
original_p_loss <- original$risky[original$condition == "loss"] / original$n[original$condition == "loss"]
```

::: {.callout-note appearance="simple" icon=false}
## Project links

- **Original paper:** Tversky, A., & Kahneman, D. (1981). The framing of decisions and the
  psychology of choice. *Science, 211*(4481), 453–458. <https://doi.org/10.1126/science.7455683>
- **Preregistration:** [OSF registration](https://osf.io/XXXXX) *(registered YYYY-MM-DD, before final data collection)*
- **Live experiment:** <https://psych251.github.io/your-repo/> (take the study exactly as participants did)
- **Repository:** <https://github.com/psych251/your-repo> (experiment code, data, and this report)
:::

## Introduction

::: {.callout-tip icon=false}
## Guidance

One or two paragraphs, no abstract and no literature review: that is in the original paper.
Say what the original study found, which specific finding you are replicating, and why it
matters to you. Write the Introduction and Methods in the past tense, even though you write
them before collecting data.
:::

## Methods

### Power analysis

::: {.callout-tip icon=false}
## Guidance

Report the original effect size, the sample needed for 80%, 90% and 95% power to detect it,
and the sample you chose and why. The table below is computed from the original study's
numbers entered at the top of this file.

Original effects are usually overestimates, so powering to the original effect tends to
produce an underpowered replication. One widely used alternative is to collect 2.5 times the
original sample, which gives 80% power to detect an effect the original study had only 33%
power to detect (Simonsohn, 2015; Experimentology §10.4). Say which approach you used, and
what the course budget allowed.
:::

```{r}
#| label: power
power_rows <- map_dfr(c(0.80, 0.90, 0.95), function(pw) {
  n <- power.prop.test(p1 = original_p_gain, p2 = original_p_loss,
                       power = pw, sig.level = 0.05)$n
  tibble(power = scales::percent(pw), n_per_group = ceiling(n), n_total = 2 * ceiling(n))
})

original_effect <- original_p_loss - original_p_gain
two_and_half <- ceiling(2.5 * original$n)

knitr::kable(power_rows, col.names = c("Power", "N per frame", "N total"),
             caption = sprintf(
               "Sample needed to detect the original effect (risky choice: %.0f%% gain vs %.0f%% loss, a difference of %.0f points), two-sided alpha = .05.",
               100 * original_p_gain, 100 * original_p_loss, 100 * original_effect))
```

For comparison, 2.5 times the original sample would be `r two_and_half[1]` in the gain frame
and `r two_and_half[2]` in the loss frame (`r sum(two_and_half)` in total).

### Planned sample

::: {.callout-tip icon=false}
## Guidance

Planned sample size and stopping rule, sampling frame (for example, "US residents on Prolific,
fluent in English, 18 or older"), known demographics, and any preselection or screening. The
stopping rule should be data-independent: "collect until N participants pass the exclusion
criteria below" rather than "until the result is clear".
:::

### Materials

::: {.callout-tip icon=false}
## Guidance

All materials. You can quote the original article directly: put the text in quotation marks
and note that it was followed precisely, or quote it and point out the exceptions. Link to the
live experiment above rather than pasting screenshots of every screen.
:::

### Procedure

::: {.callout-tip icon=false}
## Guidance

As for Materials: quote the original where you can and point out the exceptions. Describe the
sequence a participant experiences, from consent to debrief, and how long it takes.
:::

### Analysis plan

::: {.callout-tip icon=false}
## Guidance

The analysis strategy, as close to the original as possible: data cleaning rules, exclusion
rules, covariates, and the model or test. Original papers often do not spell this out; say
where you had to make a choice and what you chose.

**State the key analysis of interest explicitly.** That is the single test your replication
stands or falls on, and it must match your preregistration. You can pre-specify additional
analyses here too.
:::

**Key analysis of interest.** The proportion of participants choosing the risky program is
higher in the loss frame than in the gain frame, tested with a chi-square test of independence
on the 2 × 2 table of frame by choice. The effect size is the difference in the proportion
choosing the risky program (loss minus gain), with a 95% confidence interval.

**Exclusions.** Participants are excluded if they did not complete the study, if they answered
no speeded trial at all (every lexical-decision trial timed out, the signature of a device
without a keyboard), or if they answered fewer than 6 of the 8 lexical-decision trials
correctly.

### Differences from the original study

::: {.callout-tip icon=false}
## Guidance

Every known difference in sample, setting, procedure, materials, and analysis, one per row.
Minimizing differences is the goal, but some always remain. For each one, say whether it is
expected to matter, based on claims in the original article or later research on when the
effect does and does not appear. The rows below are the demo's real differences and show the
level of detail expected; replace them with yours.
:::

| Aspect | Original study | This replication | Expected to matter? |
|---|---|---|---|
| Sample | University students in Stanford and British Columbia classrooms | Adults recruited on Prolific | Possibly. Framing effects have been found in both student and general samples, but their size varies across populations. |
| Setting | Paper questionnaire in class | Online, unsupervised, on the participant's own computer | Unlikely for a single vignette, though attention may be lower online. |
| Wording | "an unusual Asian disease" | "an unusual disease" | Unlikely. The ethnonym is incidental to the manipulation and dated. |
| Design | Between subjects: each participant saw one frame | Between subjects, randomly assigned | No. |
| Response | Circle a program on paper | Click a button | Unlikely. |
| Other tasks | Problems presented alongside other questionnaire items | Followed by a short lexical-decision task used as an attention check | Unlikely: the check comes after the choice. |

### Methods addendum (post data collection)

::: {.callout-tip icon=false}
## Guidance

Leave this section out of the version you preregister. After data collection, report the
actual sample and any departure from the plan above.
:::

#### Actual sample

```{r}
#| label: no-data-notice
#| echo: false
#| results: asis
if (!have_data) {
  cat(
    "::: {.callout-warning icon=false}\n",
    "## No data yet\n\n",
    "There are no exported data for `", experiment_id, "` yet, so the sample and results ",
    "sections are skipped. This is expected before data collection. Export with ",
    "`npm run export -- --experiment ", experiment_id, " --key <path to your key>` and render again. ",
    "If you have exported, check that `experiment_id` at the top of this file matches ",
    "`EXPERIMENT.id` in `experiment.js`.\n",
    ":::\n", sep = "")
}
```

```{r}
#| label: load-and-exclude
#| eval: !expr have_data
participants <- read_csv(file.path(raw_dir, "participants.csv"), show_col_types = FALSE)
trials       <- read_csv(file.path(raw_dir, "trials.csv"), show_col_types = FALSE)
errors_path  <- file.path(raw_dir, "errors.csv")
errors <- if (file.exists(errors_path)) read_csv(errors_path, show_col_types = FALSE) else tibble()

complete <- participants |> filter(completed)

# Exclusion rule: replace with the rule you preregistered.
ld_accuracy <- trials |>
  filter(task == "lexical_decision") |>
  group_by(participant_id) |>
  summarise(
    ld_acc           = mean(as.logical(correct), na.rm = TRUE),
    ld_all_timed_out = all(as.logical(timed_out)),
    .groups = "drop"
  )

keep <- complete |>
  left_join(ld_accuracy, by = "participant_id") |>
  filter(!is.na(ld_acc), !ld_all_timed_out, ld_acc >= 0.75)

n_no_response <- complete |>
  left_join(ld_accuracy, by = "participant_id") |>
  filter(ld_all_timed_out %in% TRUE) |>
  nrow()
```

```{r}
#| label: funnel
#| eval: !expr have_data
funnel_levels <- c("Started", "Completed", "Passed exclusion criteria")
bind_rows(
  participants |> mutate(step = "Started"),
  complete     |> mutate(step = "Completed"),
  keep         |> mutate(step = "Passed exclusion criteria")
) |>
  mutate(step = factor(step, levels = funnel_levels)) |>
  count(step, condition, .drop = FALSE) |>
  pivot_wider(names_from = condition, values_from = n, values_fill = 0) |>
  mutate(Total = rowSums(across(-step))) |>
  rename(Step = step) |>
  knitr::kable(caption = "Participants at each step, by frame. Differences between frames in attrition or exclusion are a threat to validity and should be discussed.")
```

```{r}
#| label: sample-text
#| eval: !expr have_data
#| echo: false
#| results: asis
cat(sprintf("Of %d participants who started, %d completed the study and %d met the exclusion criteria.",
            nrow(participants), nrow(complete), nrow(keep)))
if (n_no_response > 0) {
  cat(sprintf(" %d completed session(s) contained no response to any speeded trial and were excluded on that basis.",
              n_no_response))
}
if (nrow(keep) == 0) cat(" No participants remain for analysis yet, so the results below are empty.")
cat("\n")
```

```{r}
#| label: demographics
#| eval: !expr have_data
demo <- trials |>
  filter(task == "demographics", participant_id %in% keep$participant_id) |>
  mutate(resp = map(response, ~ fromJSON(.x))) |>
  transmute(
    participant_id,
    age    = map_dbl(resp, ~ as.numeric(.x$age %||% NA)),
    gender = map_chr(resp, ~ as.character(.x$gender %||% NA))
  )

if (nrow(demo) > 0) {
  demo |>
    summarise(N = n(), `Mean age` = round(mean(age, na.rm = TRUE), 1),
              `SD age` = round(sd(age, na.rm = TRUE), 1),
              `Age not given` = sum(is.na(age))) |>
    knitr::kable(caption = "Age of the analysed sample.")
}
```

#### Differences from the pre-data-collection plan

::: {.callout-tip icon=false}
## Guidance

Any difference from the plan above, or "None". Changes made after seeing data belong here even
if they seem minor, together with the reason.
:::

## Results

### Data preparation

::: {.callout-tip icon=false}
## Guidance

Data preparation follows the analysis plan. The code that loads the exported data and applies
the exclusions is in the *Actual sample* section above; open it with "Show code". Describe
anything you did to the data beyond that.
:::

### Confirmatory analysis

```{r}
#| label: confirmatory
#| eval: !expr have_data
framing <- trials |>
  filter(task == "framing", participant_id %in% keep$participant_id) |>
  select(participant_id, condition, choice)

risky <- framing |>
  group_by(condition) |>
  summarise(risky = sum(choice == "risky"), n = n(), .groups = "drop")

tab <- table(Frame = framing$condition, Choice = framing$choice)
have_both <- all(c("gain", "loss") %in% risky$condition)

if (all(dim(tab) == c(2, 2))) {
  # No continuity correction, so this test and the effect-size interval below agree: the
  # uncorrected 2 x 2 chi-square is the same test as the two-proportion comparison.
  chi <- chisq.test(tab, correct = FALSE)
  print(tab)
  print(chi)
} else {
  message("Chi-square not run: the frame-by-choice table is ", paste(dim(tab), collapse = " x "),
          ". It needs both frames and both choices; with ", nrow(framing),
          " participant(s) in the analysis set this is expected.")
}
```

### Replication assessment

::: {.callout-tip icon=false}
## Guidance

Judge the replication on effect sizes, not only on whether the key test is significant. Compare
the replication estimate and its interval with the original: does the interval exclude zero, and
does it include the original estimate? A significant result much smaller than the original, or
a non-significant result whose interval still includes the original effect, are both
informative and should be reported as such (Experimentology §3.2). Replace the plot's demo data
with your original study's values at the top of this file.
:::

```{r}
#| label: comparison
#| eval: !expr have_data
diff_ci <- function(x_loss, n_loss, x_gain, n_gain) {
  pt <- prop.test(c(x_loss, x_gain), c(n_loss, n_gain), correct = FALSE)
  tibble(estimate = unname(pt$estimate[1] - pt$estimate[2]),
         ci_low = pt$conf.int[1], ci_high = pt$conf.int[2], p = pt$p.value)
}

orig <- diff_ci(original$risky[original$condition == "loss"], original$n[original$condition == "loss"],
                original$risky[original$condition == "gain"], original$n[original$condition == "gain"]) |>
  mutate(Study = "Original", N = sum(original$n))

if (have_both && all(risky$n > 0)) {
  rep <- diff_ci(risky$risky[risky$condition == "loss"], risky$n[risky$condition == "loss"],
                 risky$risky[risky$condition == "gain"], risky$n[risky$condition == "gain"]) |>
    mutate(Study = "Replication", N = sum(risky$n))
  comparison <- bind_rows(orig, rep)

  comparison |>
    transmute(Study, N,
              `Difference (loss − gain)` = sprintf("%.0f points", 100 * estimate),
              `95% CI` = sprintf("[%.0f, %.0f]", 100 * ci_low, 100 * ci_high),
              p = format.pval(p, digits = 2, eps = 0.001)) |>
    knitr::kable(caption = "Effect size: difference in the percentage choosing the risky program.")
} else {
  message("Replication effect size not computed yet: need at least one analysed participant in each frame.")
  comparison <- orig
}
```

```{r}
#| label: comparison-verdict
#| eval: !expr have_data
#| echo: false
#| results: asis
if (exists("rep")) {
  excludes_zero <- rep$ci_low > 0 | rep$ci_high < 0
  includes_orig <- orig$estimate >= rep$ci_low & orig$estimate <= rep$ci_high
  cat(sprintf(
    "The replication's 95%% interval %s zero and %s the original estimate of %.0f points. ",
    if (excludes_zero) "excludes" else "includes",
    if (includes_orig) "includes" else "does not include",
    100 * orig$estimate))
  cat("These two facts are the starting point for the summary in the Discussion, not the conclusion.\n")
}
```

```{r}
#| label: fig-comparison
#| eval: !expr have_data
#| fig-cap: "Percentage choosing the risky program in each frame, original study and replication. Error bars are 95% confidence intervals."
#| fig-width: 7
#| fig-height: 4
prop_ci <- function(x, n) {
  ci <- prop.test(x, n, correct = FALSE)$conf.int
  tibble(p = x / n, low = ci[1], high = ci[2])
}
plot_df <- bind_rows(
  original |> mutate(Study = "Original"),
  if (nrow(risky) > 0) risky |> mutate(Study = "Replication")
) |>
  rowwise() |>
  mutate(prop_ci(risky, n)) |>
  ungroup() |>
  mutate(Study = factor(Study, levels = c("Original", "Replication")),
         label = sprintf("n = %d", n))

ggplot(plot_df, aes(x = condition, y = p, fill = Study)) +
  geom_col(position = position_dodge(width = 0.75), width = 0.7) +
  geom_errorbar(aes(ymin = low, ymax = high), position = position_dodge(width = 0.75), width = 0.2) +
  geom_text(aes(y = 0.02, label = label), position = position_dodge(width = 0.75),
            vjust = 0, size = 3, colour = "white") +
  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
  scale_fill_manual(values = c(Original = "grey55", Replication = "#1E6E76")) +
  labs(x = "Frame", y = "Chose the risky program", fill = NULL) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top")
```

### Exploratory analyses

::: {.callout-tip icon=false}
## Guidance

Any follow-up analyses (optional). Label them clearly as exploratory: they were not planned, so
they generate hypotheses rather than test them.
:::

```{r}
#| label: exploratory-rt
#| eval: !expr have_data
ld <- trials |>
  filter(task == "lexical_decision", participant_id %in% keep$participant_id) |>
  mutate(correct = as.logical(correct), is_word = as.logical(is_word), rt = as.numeric(rt)) |>
  filter(correct, !is.na(rt))

if (nrow(ld) > 0) {
  ld |>
    group_by(`Real word` = is_word) |>
    summarise(`Mean RT (ms)` = round(mean(rt)), `SD (ms)` = round(sd(rt)), Trials = n(), .groups = "drop") |>
    knitr::kable(caption = "Exploratory: lexical-decision response times, correct trials only.")
} else {
  message("No correct lexical-decision trials with a response time yet.")
}
```

```{r}
#| label: data-quality
#| eval: !expr have_data
if (nrow(errors) > 0) {
  errors |> select(any_of(c("at", "trial_index", "message"))) |> head(20) |>
    knitr::kable(caption = "JavaScript errors reported from participants' browsers.")
} else {
  message("No errors were reported from participants' browsers.")
}
```

## Discussion

### Summary of replication attempt

::: {.callout-tip icon=false}
## Guidance

Open with one paragraph summarizing the result of the key analysis and your assessment of
whether the original finding replicated, partially replicated, or failed to replicate. Base
that judgment on the replication assessment above: the size of the effect relative to the
original, not just the p value.
:::

### Commentary

::: {.callout-tip icon=false}
## Guidance

Open-ended commentary, none of it long: (a) what the exploratory analyses suggest; (b) what the
replication means, for example whether the differences between your study and the original are
definitely, plausibly, or unlikely to be moderators of the result; and (c) any objections or
challenges raised by you or the original authors.
:::

## Data and code availability

::: {.callout-tip icon=false}
## Guidance

Edit the statement below to match what you are actually sharing. Anonymized data and all code
should be public; if something cannot be shared, say what and why.
:::

The experiment code, the anonymized data, and the code that produces every number in this
report are available at <https://github.com/psych251/your-repo>. The experiment itself can be
taken at <https://psych251.github.io/your-repo/>. The data contain no names, email addresses,
IP addresses, or Prolific IDs; the file linking participants to their Prolific IDs is kept
privately for payment and is not shared.

## AI use statement

::: {.callout-tip icon=false}
## Guidance

List every AI tool you used on this project, what you used it for, and how you checked what it
produced. Include coding assistants (for example Claude Code, Codex, Copilot), chat assistants,
and writing or editing tools. Name the model and version where you know it. "How you checked"
is the important column: say what you verified yourself, not that the output "looked right".
If you used no AI tools, replace the table with a sentence saying so.
:::

| Tool and model | What it was used for | How you checked the output |
|---|---|---|
| *e.g., Claude Code (Claude Opus 5)* | *Implemented the experiment from the original Methods section* | *Ran the automated tests; took the study myself; compared every screen with the original materials* |
| *e.g., Claude Code (Claude Opus 5)* | *Drafted the analysis code* | *Checked the exclusion counts by hand; confirmed the key test matches the preregistration* |
| *e.g., ChatGPT* | *Suggested edits to the Introduction* | *Rewrote the text myself and checked every citation against the source* |

AI tools were not used to generate, alter, or impute any participant data. I take full
responsibility for the content of this report, including any errors introduced by AI tools.
